---
title: "Sandbox"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sandbox}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(canadacovidshiny)
library(canadacovid)
library(dplyr)
```

This vignette will serve as a sandbox/notebook for working on the back-end implementation of the dashboard -- the "RMarkdown first" approach as suggested by the [Engineering Shiny](https://engineering-shiny.org/building-ispum-app.html#proto-rmdfirst) book.

## Get summaries

The first and simplest data I want to include is the latest daily counts, both overall and by province, which I retrieve with `get_summary()`

```{r}
summary_overall <- get_summary()
summary_province <- get_summary(split = "province")
glimpse(summary_province)
```


## Checking for new reports

I want to keep the API requests down to a minimum to avoid 429 errors (too many requests in a short time).

The `provinces` table lets me know if and when data has been reported:

```{r}
provinces <- canadacovid::get_provinces()
glimpse(provinces)
```

It will be helpful to put this into an easily-accessible list like so:

```{r}
province_status <- provinces %>%
  select(code, name, data_status, updated_at) %>%
  split(.$code)
str(province_status[1])
```

Then I can check the status of each province:

```{r}
province_status$NS$data_status
```
 
If the data has been reported for the day, then the data can be updated by calling the API.

## Storing the data

A package I've been meaning to try is [`pin`](https://pins.rstudio.com/).
It allows me to store R objects remotely (on *boards*), and retrieve and update that data when necessary.
Create a temporary board (that will be deleted once the R session ends):

```{r}
library(pins)

board <- board_temp()
board
```

Then, save the data to the board:

```{r}
board %>% pin_write(summary_overall, "summary_overall", type = "rds")
```

Then retrieve it:

```{r}
board %>% pin_read("summary_overall")
```

I can also get some metadata about the data, like when it was `created`:

```{r}
board %>% pin_meta("summary_overall")
```

I have a Google Cloud Platform account (free trial for now).
I'll make a `canadacovidshiny_bucket` just to try it out.
Unfortunately, I have to use the legacy `pins` API for this task, as [GCP hasn't been implemented in the modern API](https://pins.rstudio.com/articles/pins-update.html#equivalents) as of me writing this:

```{r}
gcp_board <- board_register_gcloud(
  name = "gcloud", bucket = "canadacovidshiny_bucket",
  # TCP token stored locally in my interactive session
  token = Sys.getenv("GCLOUD_ACCESS_TOKEN")
)
pin(summary_overall, name = "summary_overall", board = "gcloud")
```

```{r}
pin_get("summary_overall", board = "gcloud")
```

